{\bf Landslide: Systematic Dynamic Race Detection in Kernel-space}

### 2012-05-09 defence/defence.tex
minor update defence
update presentation

\usecolortheme[named=ForestGreen]{structure}

@@

% slide: outline
\begin{frame}{Outline}
	\textbf{Motivation: Race condition debugging}
	\begin{itemize}
		\item Systematic exploration versus stress testing
		\item Challenges of kernel-space
	\end{itemize}
	{\bf Landslide: Making it happen}
	\begin{itemize}
		\item Design and interface 
		\item Addressing challenges
	\end{itemize}
	{\bf Evaluation}
	\begin{itemize}
		\item Student user study
		\item Case study
	\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Motivation}

\subsection{Race Conditions and the Finding Thereof}

@@

\begin{frame}{Testing Techniques} % slide: sys ex vs stress testing
	\textbf{Stress Testing}: Common testing approach
	\begin{itemize}
		\item Attempting to exercise as many interleavings as possible
		\item Exposes race conditions at random
		\begin{itemize}
			\item ``If a preemption occurs at {\em just the right time}\dots''
		\end{itemize}
		\item Cryptic panic messages or machine reboots
	\end{itemize}
	\linegap

	{\bf Systematic Exploration}\related{Flanagan '05}
	\begin{itemize}
		\item Make educated guesses about when to preempt
		\item Run {\em every single} interleaving
		\item Provide better debugging information
	\end{itemize}
\end{frame}

@@

% slide: 15-410 6-week project why we need it in kernels
\begin{frame}{Finding Races in Kernel-Space}
	% cite datacollider etc?
\end{frame}

\subsection{Challenges of Kernel-space}

@@

	\linegap
	Context switcher
	\begin{itemize}
		\item Interleaving threads is no longer ``instantaneous''.
	\end{itemize}
	Runqueue tracking
	\begin{itemize}
		\item Which threads are runnable?
	\end{itemize}
	Thread lifecycle tracking
	\begin{itemize}
		\item When do threads created/destroyed?
	\end{itemize}
\end{frame}

% slide: ad-hoc communication, memory conflicts (freeing, runqueue)
\begin{frame}{Ad-Hoc Thread Communication}
	No message-passing % TODO this slide sucks

	User-space systematic exploration relies on it\related{Simsa '10}; we cannot.

	Kernel threads share state via globals, heap.
\end{frame}

\begin{frame}{The Kernel as One Program} % slide: the kernel is too big to test all at once
	\includegraphics[width=\textwidth]{pebbles.png}
	% Say: "we don't want to have to explore every possible interaction between each of these components"
\end{frame}

\begin{frame}{Contribution} % slide: summary: "sys ex is a tool \ldots``
	Systematic exploration is a {\em tool} for testing concurrent systems.
	\begin{itemize}
		\item How can it be applied in kernel-space?
		\item What should/must we ask of the user to make it possible?
	\end{itemize}
	\linegap

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Landslide}

@@

\subsection{Anatomy}
% slide: pebbles, 15-410, simics
% slide: visualisation (here address challenge #1)

\subsection{Design}
% slide: using -- annotations & in-landslide instrumentation
% slide: POR -- mostly invisible, but (transition)

\subsection{Addressing Challenges} % rename
% slide: POR, modified - with runqueue conflicts, will get no pruning
% slide: ignoring shared memory accesses - venn diagram
% slide: other_prob: decide on mutex_lock as opposed to everywhere, but...
% slide: manual shaping - have *user* specify what to focus on (automating intuition)
% slide: exciting teaser "i think this is not too far off!"

@@

\begin{frame}{Working with Students} % slide: worked with students - give some time breakdown
	Solicited students in 15-410 to use Landslide on their kernels.
	\begin{itemize}
		\item Can Landslide find bugs ``in the wild''?
		\item How much time does it take to get it working?
	\end{itemize}
	Five groups participated; four got it to work.
	\begin{itemize}
		\item Average instrumentation time 100 minutes (min 50 max 150)
		\item Average customisation time 25 minutes (min 0 max 60)
	\end{itemize}
\end{frame}

\begin{frame}{Found Bugs} % slide: anecdotes: two deterministic bugs & two races
	All four groups who got it to work found bugs.

@@

\begin{frame}{Studying Bugs In-Depth} % slide: "I have built a software smarter than me!" 30-40sec to find
	Worked with two kernels: my own, and one I graded

	``I have made something smarter than me.''
	\begin{itemize}
		\item \texttt{double\_wait} - two parent threads \texttt{wait()} on one exiting child
		\item Expected: One thread reaps child, the other returns failure
		\item Got: \texttt{failed assertion `nobe->cond\_queue\_next == NULL'}
	\end{itemize}
\end{frame}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Summary}

% slide: recap challenges and how we dealt with them
% slide: conclusion
\begin{frame}{Conclusion}
	\textbf{Systematic exploration in kernel-space}
	% TODO
\end{frame}

@@

\end{document}
%\begin{frame}{Partial Order Reduction}
%	\begin{center}
%	The next 6 slides won't be on the exam.
%
%		\Huge \dbend
%	\end{center}
%\end{frame}
%
%\begin{frame}{Partial Order Reduction}
%	A {\bf transition} is\ldots
%	\begin{itemize}
%		\item The sequence of instructions executed between choice points.
%	\end{itemize}
%	To express concurrency between transitions:
%	% Say: As a prerequisite for this algorithm, we need to build two sets to express the concurrency relations between the transitions
%	\begin{itemize}
%		\item Independence relation
%		\item Happens-before relation
%	\end{itemize}
%	% Say: This is hard, and I don't expect you to understand it during this lecture; you won't be tested on it or anything. My goal in presenting it is to convey how intensely cool the algorithm is, and hopefully you will keep thinking about it later. I'd be happy to answer questions about it later.
%\end{frame}
%
%\begin{frame}[fragile]{Partial Order Reduction - Independence Relation}
%	\begin{columns}[l]
%		\column{0.5\textwidth}
%		\begin{center}
%		\begin{verbatim}
%		void thread_A() {
%		    foo = 42;
%		}
%		void thread_B() {
%		    if (bar == 1337)
%		        print "Hello"
%		}
%		\end{verbatim}
%		\end{center}
%		\linegap
%
%		$A$ and $B$ are {\bf independent}.
%		\column{0.5\textwidth}
%		\begin{center}
%		\begin{verbatim}
%		void thread_A() {
%		    foo = 42;
%		}
%		void thread_C() {
%		    if (foo == 1337)
%		        print "Hello"
%		}
%		\end{verbatim}
%		\end{center}
%		\linegap
%
%		$A$ and $C$ {\bf conflict}.
%	\end{columns}
%\end{frame}
%
%\begin{frame}{Partial Order Reduction - Independence Relation}
%	Consider a trace of all shared memory reads/writes
%
%	\linegap
%	Two transitions (instruction sequences) $A$ and $B$ are independent iff:
%	\begin{itemize}
%		\item $Writes(A) \cap Writes(B) = \emptyset$
%		\item $Writes(A) \cap Reads(B) = \emptyset$
%		\item $Reads(A) \cap Writes(B) = \emptyset$
%	\end{itemize}
%\end{frame}
%
%\begin{frame}{Partial Order Reduction - Happens-before Relation}
%	\begin{columns}[l]
%		\column{0.5\textwidth}
%		%\includegraphics[width=\textwidth]{hb.png}
%		\column{0.5\textwidth}
%		Is a transition ``enabled'' by a previous one?
%		\begin{itemize}
%			\item $A$ enables $B$
%			\item $B$ enables $C$ (same thread)
%			\item $B$ enables $D$
%		\end{itemize}
%%		Transitive closure
%%		\begin{itemize}
%%			\item $HB^* = \{AB, AC, AD, BC, BD\}$
%%		\end{itemize}
%		$C$ and $D$ are {\bf concurrent}!
%	\end{columns}
%\end{frame}
%
%\begin{frame}{Partial Order Reduction - Tagging}
%	Run POR at the end of each branch
%	\begin{itemize}
%		\item Identify which branches to explore next
%		\item Tag only a subset of all branches (hopefully)
%	\end{itemize}
%\end{frame}
%
%\begin{frame}{Partial Order Reduction - Tagging}
%	For each transition $T$, find an {\bf evil ancestor} $E$:
%	\begin{itemize}
%		\item $E$ and $T$ are concurrent.
%		\item $E$ and $T$ are not independent.
%		\item Intuitively: $T$ and $E$ could be reordered, and might behave differently.
%	\end{itemize}
%	Find a {\bf good sibling} $G$ of $E$:
%	\begin{itemize}
%		\item $G$ is the same thread as $T$
%		% Say: (if no such sibling, all siblings)
%		\item Intuitively: Find a branch in which $T$ runs before $E$.
%	\end{itemize}
%	Explore tagged branches depth-first.
%\end{frame}
%
%\begin{frame}{Partial Order Reduction}
%	\begin{center}
%		{\Huge \dbend}
%		\linegap
%		\linegap
%		\linegap
%
%		Feel free to ask me about this later.
%	\end{center}
%\end{frame}

>>>>

\usecolortheme[named=RoyalBlue]{structure}

@@

% slide: outline
\begin{frame}{Outline}
	\textbf{Motivation: Concurrency debugging}
	\begin{itemize}
		\item Systematic exploration versus stress testing
		\item Challenges of kernel-space
	\end{itemize}
	\linegap

	{\bf Tool: Landslide}
	\begin{itemize}
		\item Design and interface 
		\item Addressing challenges
	\end{itemize}
	\linegap

	{\bf Evaluation: Finding Races}
	\begin{itemize}
		\item Student user study
		\item Case study
	\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Motivation}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Race Conditions and the Finding Thereof}

@@

\begin{frame}{Testing Techniques} % slide: sys ex vs stress testing
	\textbf{Stress Testing:} Common testing approach
	\begin{itemize}
		\item Attempting to exercise as many interleavings as possible
		\item Exposes race conditions at random
		\begin{itemize}
			\item ``If a preemption occurs at {\em just the right time}\dots''
		\end{itemize}
		\item Cryptic panic messages or machine reboots
	\end{itemize}
	\linegap

	{\bf Systematic Exploration} % \related{Flanagan '05}
	\begin{itemize}
		\item Make educated guesses about when to preempt
		\item Run {\em every single} interleaving
		\item Provide better debugging information
	\end{itemize}
\end{frame}

@@

\begin{frame}{Finding Races in Kernel-Space} % slide: 15-410 6-week project why we need it in kernels
	15-410: Operating System Design and Implementation
	\begin{itemize}
		\item Project 3: students write a kernel in 6 weeks
		\item ``Pebbles'': UNIX-like, fully preemptible
	\end{itemize}
	\linegap

	Previous work in kernel-space has not used systematic exploration.
	\begin{itemize}
		\item Data race detection\related{Erickson '10}
		\item Failure resiliency\related{Kadav '09}
	\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Challenges of Kernel-space}

\breakslide{Why is this environment different from all other environments?}

@@

	\linegap
	Context switcher
	\begin{itemize}
		\item Interleaving threads is no longer ``instantaneous''.
	\end{itemize}
	Runqueue tracking
	\begin{itemize}
		\item Which threads are runnable?
	\end{itemize}
	Thread lifecycle tracking
	\begin{itemize}
		\item When are threads created/destroyed?
	\end{itemize}
\end{frame}

% slide: ad-hoc communication, memory conflicts (freeing, runqueue)
%\begin{frame}{Ad-Hoc Thread Communication}
%	No message-passing % TODO this slide sucks
%	
%	User-space systematic exploration relies on it\related{Simsa '10}; we cannot.
%	
%	Kernel threads share state via globals, heap.
%\end{frame}

\begin{frame}{The Kernel as One Program} % slide: the kernel is too big to test all at once
	\begin{itemize}
		\item ``Everything interleaves with everything else''?
		\item What is a ``bug'', anyway?\related{Baumann '09}
	\end{itemize}
	\begin{center}
	\includegraphics[width=\textwidth]{pebbles.png}
	\end{center}
	% Say: "we don't want to have to explore every possible interaction between each of these components"
\end{frame}

\begin{frame}{Contribution} % slide: summary: "sys ex is a tool \ldots``
	Systematic exploration is a {\em tool} for testing concurrent systems.
	\begin{itemize}
		\item How can it be applied in kernel-space?
		\item What can we ask of the user to make it possible?
	\end{itemize}
	\linegap

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Landslide}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Design}

% slide: pebbles, 15-410, simics
\begin{frame}{Environment}
\end{frame}

% slide: visualisation (here address challenge #1)
\begin{frame}{Anatomy of Landslide}
	\begin{center}
	\includegraphics[width=0.7\textwidth]{landslide.png}
	\end{center}
\end{frame}

% slide: using -- annotations & in-landslide instrumentation
\begin{frame}{Using Landslide}
\end{frame}

% slide: POR -- mostly invisible, but (transition)
\begin{frame}{State Space Reduction}
	\begin{center}
	\includegraphics[width=\textwidth]{undiamond1.png}
	\end{center}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Addressing Challenges} % rename

\begin{frame}{Shared Memory Conflicts} % slide: ignoring shared memory accesses - venn diagram
	\begin{center}
	\includegraphics[width=\textwidth]{shm-conflict0.png}
	\end{center}
\end{frame}
\begin{frame}{Shared Memory Conflicts}
	\begin{center}
	\includegraphics[width=\textwidth]{shm-conflict1.png}
	\end{center}
\end{frame}
\begin{frame}{Shared Memory Conflicts}
	\begin{center}
	\includegraphics[width=\textwidth]{shm-conflict2.png}
	\end{center}
\end{frame}
\begin{frame}{Shared Memory Conflicts}
	\textbf{Problem:} Many conflicts we don't care about
	\begin{itemize}
		\item Every transition accesses scheduler state
		\item Many transitions access ``unrelated'' global objects
	\end{itemize}
	\linegap

	{\bf Solution:} Only consider ``relevant'' memory conflicts
	\begin{itemize}
		\item Ignore scheduler, global objects ``known to be correct''
		\begin{itemize}
			\item Sacrifice ability to test these to efficiently test everything else.
		\end{itemize}
	\end{itemize}
\end{frame}
% slide: identifying bugs
\begin{frame}{Identifying Bugs}
	\textbf{How do we know we've found a bug?}
\end{frame}
% slide: other_prob: decide on mutex_lock as opposed to everywhere, but...
% slide: manual shaping - have *user* specify what to focus on (automating intuition)
\begin{frame}{Limiting Decision Points}
\end{frame}
% slide: exciting teaser "i think this is not too far off!"
\begin{frame}{Teaser}
\end{frame}

@@

\begin{frame}{Working with Students} % slide: worked with students - give some time breakdown
	Solicited students in 15-410 to use Landslide on their kernels.
	\begin{itemize}
		\item Can Landslide find bugs ``in the wild''?
		\item How much time does it take to get it working?
	\end{itemize}
	\linegap

	Five groups participated; four got it to work.
	\begin{itemize}
		\item Average instrumentation time 100 minutes (50 to 150)
		\item Average customisation time 25 minutes (0 to 60)
	\end{itemize}
\end{frame}

\begin{frame}{Landslide Victories} % slide: anecdotes: two deterministic bugs & two races
	All four groups who got it to work found bugs.
	\linegap

@@

\begin{frame}{Studying Bugs In-Depth} % slide: "I have built a software smarter than me!" 30-40sec to find
	Worked with two kernels: my own, and one I graded
	\linegap

	``I have made something smarter than me.''
	% Heh, if it were really smarter than me, I wouldn't have had to debug it.
	\begin{itemize}
		\item \texttt{double\_wait} - two parent threads \texttt{wait()} on one exiting child
		\item Expected: One thread reaps child, the other returns failure
		\item Got: \texttt{failed assertion `nobe->cond\_queue\_next == NULL'}
	\end{itemize}
\end{frame}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusion}

\subsection{Future Work}
\begin{frame}{Promising Directions}
	Top 15-410 students, effectively supervised, can find hard bugs faster
	\begin{itemize}
		\item Use as a debugging tool may help students learn more
	\end{itemize}
	\linegap

	After Pebbles\dots
	\begin{itemize}
		\item Potential application to production kernels
		\begin{itemize}
			\item General purpose OSes, device drivers, embedded kernels
			% Be ready to field questions about SMP.
		\end{itemize}
	\end{itemize}
	\linegap

	Emphasis on ``steering'' of test parameters by experts
	\begin{itemize}
		\item Perhaps we can automate this steering?
	\end{itemize}
	\linegap
\end{frame}

\subsection{Summary}
\begin{frame}{Summary}
	\textbf{Systematic exploration in kernel-space}
	\begin{itemize}
		\item Use internal kernel abstractions to understand concurrency behaviour
		\item Relying on user's knowledge makes testing easier
	\end{itemize}
	\linegap
	\textbf{Landslide}
	\begin{itemize}
		\item After polishing, a useful tool for 15-410
		\item A first step towards sophisticated kernel debugging techniques
	\end{itemize}

	\linegap
	\linegap
	\begin{center}
	{\scriptsize \em Please hold on. This transit is approaching the Landslide terminal, \\baggage claim, ground transportation, and ticketing.}
	\end{center}
\end{frame}

% slide: bonus slides

\end{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%\begin{frame}{Partial Order Reduction}
%	\begin{center}
%	The next 6 slides won't be on the exam.
%
%		\Huge \dbend
%	\end{center}
%\end{frame}
%
%\begin{frame}{Partial Order Reduction}
%	A {\bf transition} is\ldots
%	\begin{itemize}
%		\item The sequence of instructions executed between choice points.
%	\end{itemize}
%	To express concurrency between transitions:
%	% Say: As a prerequisite for this algorithm, we need to build two sets to express the concurrency relations between the transitions
%	\begin{itemize}
%		\item Independence relation
%		\item Happens-before relation
%	\end{itemize}
%	% Say: This is hard, and I don't expect you to understand it during this lecture; you won't be tested on it or anything. My goal in presenting it is to convey how intensely cool the algorithm is, and hopefully you will keep thinking about it later. I'd be happy to answer questions about it later.
%\end{frame}
%
%\begin{frame}[fragile]{Partial Order Reduction - Independence Relation}
%	\begin{columns}[l]
%		\column{0.5\textwidth}
%		\begin{center}
%		\begin{verbatim}
%		void thread_A() {
%		    foo = 42;
%		}
%		void thread_B() {
%		    if (bar == 1337)
%		        print "Hello"
%		}
%		\end{verbatim}
%		\end{center}
%		\linegap
%
%		$A$ and $B$ are {\bf independent}.
%		\column{0.5\textwidth}
%		\begin{center}
%		\begin{verbatim}
%		void thread_A() {
%		    foo = 42;
%		}
%		void thread_C() {
%		    if (foo == 1337)
%		        print "Hello"
%		}
%		\end{verbatim}
%		\end{center}
%		\linegap
%
%		$A$ and $C$ {\bf conflict}.
%	\end{columns}
%\end{frame}
%
%\begin{frame}{Partial Order Reduction - Independence Relation}
%	Consider a trace of all shared memory reads/writes
%
%	\linegap
%	Two transitions (instruction sequences) $A$ and $B$ are independent iff:
%	\begin{itemize}
%		\item $Writes(A) \cap Writes(B) = \emptyset$
%		\item $Writes(A) \cap Reads(B) = \emptyset$
%		\item $Reads(A) \cap Writes(B) = \emptyset$
%	\end{itemize}
%\end{frame}
%
%\begin{frame}{Partial Order Reduction - Happens-before Relation}
%	\begin{columns}[l]
%		\column{0.5\textwidth}
%		%\includegraphics[width=\textwidth]{hb.png}
%		\column{0.5\textwidth}
%		Is a transition ``enabled'' by a previous one?
%		\begin{itemize}
%			\item $A$ enables $B$
%			\item $B$ enables $C$ (same thread)
%			\item $B$ enables $D$
%		\end{itemize}
%%		Transitive closure
%%		\begin{itemize}
%%			\item $HB^* = \{AB, AC, AD, BC, BD\}$
%%		\end{itemize}
%		$C$ and $D$ are {\bf concurrent}!
%	\end{columns}
%\end{frame}
%
%\begin{frame}{Partial Order Reduction - Tagging}
%	Run POR at the end of each branch
%	\begin{itemize}
%		\item Identify which branches to explore next
%		\item Tag only a subset of all branches (hopefully)
%	\end{itemize}
%\end{frame}
%
%\begin{frame}{Partial Order Reduction - Tagging}
%	For each transition $T$, find an {\bf evil ancestor} $E$:
%	\begin{itemize}
%		\item $E$ and $T$ are concurrent.
%		\item $E$ and $T$ are not independent.
%		\item Intuitively: $T$ and $E$ could be reordered, and might behave differently.
%	\end{itemize}
%	Find a {\bf good sibling} $G$ of $E$:
%	\begin{itemize}
%		\item $G$ is the same thread as $T$
%		% Say: (if no such sibling, all siblings)
%		\item Intuitively: Find a branch in which $T$ runs before $E$.
%	\end{itemize}
%	Explore tagged branches depth-first.
%\end{frame}
%
%\begin{frame}{Partial Order Reduction}
%	\begin{center}
%		{\Huge \dbend}
%		\linegap
%		\linegap
%		\linegap
%
%		Feel free to ask me about this later.
%	\end{center}
%\end{frame}
